%              %
%%            %%
%%% PREAMBLE %%%
%%            %%
%              %

% Document class
\documentclass[11pt]{beamer}

\usetheme{Boadilla}
\usecolortheme{beaver}
\useinnertheme{rectangles}

\setbeamertemplate{navigation symbols}{}

% Font
\usepackage{fontspec}
\setmainfont[%
	SmallCapsFont={* Caps},%	enable small capital font family
	SlantedFont={* Slanted},%	enable slanted font family
]{Latin Modern Roman}

% Language
\usepackage{polyglossia}
\setdefaultlanguage{english}
\setotherlanguage{french}

\usepackage[autostyle=true]{csquotes}

% Floats
\usepackage{float}
\usepackage{booktabs}
\usepackage{multicol}


%              %
%%            %%
%%% DOCUMENT %%%
%%            %%
%              %

% Information
\title{Installing \LaTeX}
\subtitle{Distributions and editors}
\author[A. Quenon]{Alexandre Quenon}

% Text
\begin{document}
% *** Title page *** %
\begin{frame}
	\titlepage
\end{frame}


% *** Contents *** %
\begin{frame}
	\tableofcontents
\end{frame}


% *** Tutorial *** %
\section{Software}

\begin{frame}
	\frametitle{\LaTeX-related programs}
	
	To write a document with \LaTeX{}, the user must install specific programs, as for any other text editor.
	These are:
	\begin{itemize}
		\item a distribution,
		\item an editor.
	\end{itemize}

	The \LaTeX{} distribution contains all tools necessary to generate a document (usually a PDF) from the \enquote{code}.
	The editor is a program which simplifies the \LaTeX{} writing by providing shortcuts for compilation (cf. next tutorial) and shortcuts for text styling (e.g., \texttt{CTRL+B} creates bold text).
\end{frame}


\section{Distribution}

\begin{frame}
	\frametitle{Which distribution to choose}
	\framesubtitle{Comparison of LaTeX distributions based on OS compatibility}
	
	\begin{table}
		\begin{tabular}{*{4}{c}} \toprule
			&\multicolumn{3}{c}{Distributions} \\ \cmidrule(l){2-4}
			OS			& \TeX{} Live	& MiK\TeX{}	& Mac\TeX{}	\\ \midrule
			Windows		& 	yes			& 	yes   	& 	no   	\\ 
			Linux		& 	yes			& 	yes		& 	yes		\\ 
			MacOS   	& 	no			& 	yes		& 	yes	 	\\ \bottomrule
		\end{tabular}
	\end{table}
\end{frame}

\begin{frame}
	\frametitle{Which distribution to choose}
	\framesubtitle{Feedback from my own experience}
	
	As I have used both Windows and Linux (mainly CentOS) operating systems but never Mac, I cannot provide any feedback about \alert{Mac\TeX{}}. Sorry for Mac users\dots
	
	However, I have used extensively \alert{MiK\TeX{}} (on Windows only) and \alert{\TeX{} Live} (both Linux and Windows):
	\begin{itemize}
		\item I faced problems for using some packages or compilers with MiK\TeX{} that were not existing with \TeX{} Live such as
		\begin{itemize}
			\item compilation issues when using LuaLaTeX (cf. tutorial B002),
			\item \texttt{biber} engine not recognised while I was using the \texttt{biblatex} package,
		\end{itemize}
		\item this is what I experienced
		\begin{itemize}
			\item I compared up-to-date versions,
			\item I did it a few years ago so it could have changed since the comparison was done,
		\end{itemize}
		\item I \textbf{recommend} the use of \textbf{\TeX{} Live}.
	\end{itemize}
\end{frame}


\section{Editor}

\begin{frame}
	\frametitle{Which editor to choose}
	\framesubtitle{Comparison of LaTeX editors based on OS compatibility}
	
	\begin{table}
		\begin{tabular}{*{4}{c}} \toprule
			&\multicolumn{3}{c}{Editors} \\ \cmidrule(l){2-4}
			OS			& \TeX{}studio	& \TeX{}maker	& \TeX{}nicCenter	\\ \midrule
			Windows		& 	yes			& 	yes   		& 	yes   	\\ 
			Linux		& 	yes			& 	yes			& 	no		\\ 
			MacOS   	& 	yes			& 	yes			& 	no	 	\\ \bottomrule
		\end{tabular}
	\end{table}
\end{frame}

\begin{frame}
	\frametitle{Which editor to choose}
	\framesubtitle{Feedback from my own experience}
	
	As I have used both \TeX{}maker and \TeX{}studio but \alert{\TeX{}nicCenter}.
	Sorry that I cannot provide any feedback about it\dots
	
	I have used a lot \alert{\TeX{}maker} (on Windows only) then I switched to  \alert{\TeX{}studio} (both Linux and Windows) because to my opinion:
	\begin{itemize}
		\item the tool was more \enquote{clever} considering the key management (references, bibliography),
		\item it proposes automatic alignment in \texttt{tabular} environments (even though this is not working when the \texttt{multicolumn} command is used).
	\end{itemize}
	However, for those who like a fine control, e.g. for compilation, \TeX{}maker is providing a more visible interface.
	
	Anyway, the editor is a matter of taste. So try some and do not hesitate to change and to compare them to find yours.
\end{frame}


\section{Installation examples}

\begin{frame}
	\frametitle{Installation process}
	\framesubtitle{Preliminary notes}
	
	Here after examples of installation for different distributions are shown.
	Among the details provided for each step, the installer download is intended for Windows users.
	For Linux users, it is strongly recommended to use the program installer utility related to the Linux distribution.
	For example, \texttt{apt-get} on Ubuntu.
\end{frame}

\begin{frame}
	\frametitle{Installation process}
	\framesubtitle{Installing \TeX{} Live}
	
	Here below the installation of the \TeX{} Live distribution together with the \TeX{}studio editor is presented.
	
	Follow the steps:
	\begin{enumerate}
		\item install the \TeX{} Live distribution
			\begin{enumerate}
				\item download the \href{https://www.tug.org/texlive/acquire-netinstall.html}{installer} and run it,
				\item choose \enquote{Custom install} and click on \enquote{Next},
				\item select the scheme and the packages that you need
				\begin{itemize}
					\item if you have any doubt select the \enquote{scheme-full} that installs all packages,
					\item otherwise you can change the installation collections,
				\end{itemize}
				\item click on \enquote{Install TeX Live} $\rightarrow$ this takes a moment,
			\end{enumerate}
		\item install the \TeX{}studio editor
			\begin{enumerate}
				\item download the \href{https://www.texstudio.org/}{installer} and run it,
				\item the editor should detect your \LaTeX{} distribution at first run.
			\end{enumerate}	
	\end{enumerate}
\end{frame}

\begin{frame}
	\frametitle{Installation process}
	\framesubtitle{Installing MiK\TeX{}}
	
	Here below the installation of the MiK\TeX{} distribution together with the \TeX{}studio editor is presented.
	
	Follow the steps:
	\begin{enumerate}
		\item install the MiK\TeX{} distribution
		\begin{enumerate}
			\item go in the \href{https://miktex.org/download}{download} section of the MiK\TeX{} website,
			\item click on the \enquote{All downloads} tab,
			\item select the installer that you need
			\begin{itemize}
				\item Basic Installer for a standard \LaTeX{} system,
				\item Net Installer for a complete \LaTeX{} system,
			\end{itemize}
			\item run the installer $\rightarrow$ this takes a moment,
		\end{enumerate}
		\item install the \TeX{}studio editor
		\begin{enumerate}
			\item download the \href{https://www.texstudio.org/}{installer} and run it,
			\item the editor should detect your \LaTeX{} distribution at first run.
		\end{enumerate}	
	\end{enumerate}
\end{frame}

 
\end{document}